Determinism, Complexity, and Predictability in Computer Performance

### 2014-02-27 intro.tex

Unfortunately, as a corollary of the undecidability of the halting problem: there cannot exist \emph{one} universally ideal forecasting scheme for even completely noise-free deterministic time series\cite{sfi-prediction}, let alone an arbitrary time series. This naturally leads to an important and hard question: given a complicated real-valued time series, is it possible to reliably quantify where on this spectrum the time series lies? That is, when little or nothing is known about the underlying system's generating process (e.g., linear, nonlinear, deterministic, stochastic, stationary, non-stationary etc.) how can we infer from a noisy real-valued time series if the signal is too \emph{complex} to forecast?

>>>>

Unfortunately, as a corollary of the undecidability of the halting problem: there cannot exist \emph{one} universally ideal forecasting scheme for even completely noise-free deterministic time series\cite{weigend-book}, let alone an arbitrary time series. This naturally leads to an important and hard question: given a complicated real-valued time series, is it possible to reliably quantify where on this spectrum the time series lies? That is, when little or nothing is known about the underlying system's generating process (e.g., linear, nonlinear, deterministic, stochastic, stationary, non-stationary etc.) how can we infer from a noisy real-valued time series if the signal is too \emph{complex} to forecast?
